Hydras, Ordinals \& Co. \\ A library in Coq of entertaining formal mathematics

### 2021-10-12 doc/JFLA2022/paper.tex
more refs and expansion of gaia lineage

Maintenance includes not only adaptation to changes in new proof assistant versions, but also reorganization to ease future maintenance effort and accommodate new contributions. Documentation is usually two-fold: \emph{source code comments} that describe specific definitions and results in-line and \emph{books} that carefully introduce the library and its idioms~\cite{MCB}. The former tends to be terse and dense and serves experienced users, while the latter is more long-winded and exhaustive and serves beginners. As a library changes and expands, all its documentation needs to be made consistent and complete. Authors use a wide range of techniques, including literate programming~\cite{Knuth1984} and custom tools, to pretty-print and check source code snippets and generate proof assistant output.

@@

\subsection{Vision}
The \Hydras project, part of \community~\cite{CoqCommunity} aims to be an experimental platform for the collaborative development of commented libraries of formal proofs. \community is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting \coq packages, and working collaboratively on documentation projects. The \Hydras project demonstrates that these two goals are not independent: interesting \coq packages can become the basis for new documentation.
%
This umbrella project now includes evolved versions of the former Cantor and Additions libraries (under the new names of Hydra-battles and Addition-chains), the Ackermann sub-library, extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel} and a bridge to the \gaia library (by José Grimm~\cite{Gaia,grimm:hal-00911710}).
%
By following this approach of commenting interesting \coq packages, we provide new documentation content, that contributes to the diversity of the thriving \coq documentation ecosystem.

@@

We plan to extend our libraries in two main directions.
The \gaia project~\cite{Gaia}, also maintained on \community, contains a development by José Grimm~\cite{grimm:hal-00911710},  dedicated  to the implementation in \coq (SSreflect) of books from  Bourbaki's Elements of Mathematics. It contains data structures which are compatible with our ordinal notations. 
Our plan is to build a  bridge between  the combinatorial results of \Hydras and the set-theoretic content of \gaia, and make possible the transfer of theorems between both libraries.
We aim also to write in \coq a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.

>>>>

Maintenance includes not only adaptation to changes in new proof assistant versions, but also reorganization to ease future maintenance effort and accommodate new contributions. Documentation is usually two-fold: \emph{source code comments} that describe specific definitions and results in-line and \emph{books} that carefully introduce the library and its idioms~\cite{MCB}. The former tends to be terse and dense and serves experienced users, while the latter is more long-winded and exhaustive and serves beginners. As a library changes and expands, all its documentation needs to be made consistent and complete. Authors use a wide range of techniques, including literate programming~\cite{Knuth1984} and custom tools~\cite{vanDoorn2020}, to pretty-print and check source code snippets and generate proof assistant output.

@@

\subsection{Vision}
The \Hydras project, part of \community~\cite{CoqCommunity}, aims to be an experimental platform for the collaborative development of commented libraries of formal proofs. \community is a community organization that we have founded in 2018 with two goals in mind: providing a solution for the long-term maintenance of interesting \coq packages, and working collaboratively on documentation projects. The \Hydras project demonstrates that these two goals are not independent: interesting \coq packages can become the basis for new documentation.
%
This umbrella project now includes evolved versions of the former Cantor and Additions libraries (under the new names of Hydra-battles and Addition-chains), the Ackermann sub-library, extracted from Russel O'Connor's Goedel library~\cite{OConnor05, Goedel} and a bridge to the \gaia library (by José Grimm~\cite{Gaia,grimm:hal-00911710}).
%
By following this approach of commenting interesting \coq packages, we provide new documentation content, that contributes to the diversity of the thriving \coq documentation ecosystem.

@@

We plan to extend our libraries in two main directions.

The \gaia project~\cite{Gaia}, also maintained on \community, contains a development by José Grimm~\cite{grimm:hal-00911710},  dedicated  to the implementation in \coq (SSReflect) of books from  Bourbaki's Elements of Mathematics. The set theory underpinning \gaia is derived from an earlier development by Carlos Simpson~\cite{Simpson2004,CatsZFCContrib}. \gaia contains data structures which are compatible with our ordinal notations.
Our plan is to build a  bridge between  the combinatorial results of \Hydras and the set-theoretic content of \gaia, and make possible the transfer of theorems between both libraries.
We aim also to write in \coq a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
